---
title: "CIS585 Project Report"
author: "Logan O'Connor"
date: "May 4, 2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Background

This analysis was done to determine whether or not video game developers who
publish titles to Steam can get any meaningful feedback from the reviews players
posted on the store page. This report will show the process of retrieving the reviews
with Steam's web API using the rvest and jsonlite package, text mining using the tm and hunspell packages, and finally visualizing the bigrams and their frequencies using ggwordcloud. 

###Packages used

```{r, include = TRUE}
#Required packages
require(rvest)
require(jsonlite)
require(tm)
require(ggwordcloud)
require(dplyr)
require(anytime)
require(tidytext)
require(hunspell)
```

###Setting up variables for API request
```{r, include = TRUE}
steamappid <- "289070" #ID of the game/app in Steam
url <- paste0('https://store.steampowered.com/appreviews/',steamappid,'?json=1&filter=all&language=english&day_range=9223372036854775807&num_per_page=100&purchase_type=all&start_offset=')
storeurl <- read_html(paste0("https://store.steampowered.com/app/",steamappid))
gamename <- tolower(storeurl %>% html_nodes("div.apphub_AppName") %>% html_text())
gamename <- strsplit(gamename, " ")[[1]]
```

Steam's web API returns a page of JSON based on what parameters you give it. The documention for this API request can be read [here](https://partner.steamgames.com/doc/store/getreviews). Every application in the Steam Store has it's own unique ID, the API uses this ID to tell what app we want to request reviews from. For this demonstration we picked the game "Sid Meier's Civilization VI" and stored it's ID in the steamappid variable. Next, using the paste0 function, create part of our API request URL. The parameters in the JSON request are set to return what kind of reviews are wanted, in this case it will be pulling all reviews that are english. The day range parameter is set to 9223372036854775807 due to it being the maximum amount of days it can be set to. The JSON request can only return a maximum of 100 reviews per request so the num_per_page parameter is set to match that. Finally the start_offset is set to be blank because the script will be cycling through this based on how many reviews it is told to grab. Also in this section of code, using the Rvest package, the HTML of the store page for the game is read and the name of the game is stored in a variable and is then split into a list of words for purpose of removing any instance of the games name from the review data.

###Gathering review data

```{r, include = TRUE}
pages <- list() #The list that the raw JSON will be stored in
i <- 0 #The starting offset value for the API request
amount <- 1000 #amount of reviews
numofentries <- 0

while(numofentries < amount){
  
  rawjson <- fromJSON(paste0(url,i)) #Getting JSON
  
  message("Retrieving review ", numofentries, " out of ", amount, ".")
  
  if (rawjson[["query_summary"]][["num_reviews"]] == 0) #Stopping the loop when there are no reviews left
  {
    message("Done")
    break
  }
  
  pages[[i+1]] <- rawjson$reviews #Storing JSON in list
  reviewdata_df <- rbind_pages(pages) #Turning list into a data frame
  reviewdata_df <- reviewdata_df %>% distinct(recommendationid, .keep_all = TRUE)#Removing any duplicate entries that might have snuck in
  numofentries <- length(reviewdata_df$recommendationid)
  
  i <- i+100 #Moving to next set of reviews
  
}
message("Done")
message(numofentries, " reviews with the oldest being from ", anytime(tail(reviewdata_df$timestamp_created, 1)) )
```
Here the in this while loop is where the JSON file that the Steam API returns is stored in a data frame. First before starting the loop an empty list needs to be made for the data to be stored in, after that the starting offset represented by the variable "i" is set and the amount of reviews wanting to be pulled is declared. More data is always better but this script can be quite intense on RAM. Depedning on how much RAM a person's computer has it can allow for more or less reviews to be analyzed. The computer in this demostration had 8GB of RAM and after testing I found the most reviews that can be analyzed with 8GB of RAM without the script erroring is about 3000. For this example though 1000 reviews will be used. The number of reviews the loop has grabbed so far is also tracked to compare against the number of reviews requested so the loop will stop once the number of reviews wanted is reached. Sometimes the Steam API will return duplicate reviews in a different offset so to avoid this issue after each itteration of the reviews being added to the data frame duplicate entries are removed. After the loop is finished gathering the number of reviews requested it will display a message saying how many reviews it gatherd and the time stamp of the oldest review using the anytime package. 

###Text Mining and Cleaning
```{r, include=TRUE}
#Organizing data

negreview_df <- reviewdata_df[reviewdata_df$voted_up == FALSE,] #Data frame of negative reviews
posreview_df <- reviewdata_df[reviewdata_df$voted_up == TRUE,] #Data frame of positive reviews
#-----------------------------
rm(pages, rawjson, reviewdata_df) #Removing unused data to clear up memory
```
The newly created data frame of reviews is seperated into two data frames, one for the positive reviews and one for the negative. The lists and data frames used in the retrival process are removed from the enviroment to clear memory. 

```{r, include=TRUE}
posreviews <- posreview_df$review #Getting only the review part of the data frame
negreviews <- negreview_df$review
posreviews <- iconv(posreviews, 'UTF-8', 'ASCII') #Removing emojis or other non-text characters
negreviews <- iconv(negreviews, 'UTF-8', 'ASCII')

```
A new list is made of only the part of the data frame that contains reviews. The two lists are then converted into UTF-8 character types to remove emoji and other characters that Steam's reviews allow but do not cooperate with some of text mining functions used in this script. 

```{r, include=TRUE}
# Gathering list of misspelled words that will be cleaned and combined to be used later
badpos <- hunspell(posreview_df$review) #Getting list of misspelled words to remove
badneg <- hunspell(negreview_df$review)
badpos <- unlist(badpos)
badneg <- unlist(badneg)

badpos <- iconv(badpos, 'UTF-8', 'ASCII')
badneg <- iconv(badneg, 'UTF-8', 'ASCII')

badneg <- removeNumbers(badneg)
badpos <- removeNumbers(badpos)

badpos <- removePunctuation(badpos)
badneg <- removePunctuation(badneg)

badpos <- removeWords(badpos, "")
badneg <- removeWords(badneg, "")

badpos <- unique(badpos)
badneg <- unique(badneg)

badwords <- c(badpos, badneg)

badwords <- unique(badwords)
# End of misspelled word cleaning
```
The hunspell package is used here to gather a list of words that are misspelled in each set of reviews. The list is then cleaned and combined into one list of misspelled words that apear in both sets of reviews. The words will be removed in the text cleaning process later.

```{r, include=TRUE}
posreviews <- VCorpus(VectorSource(posreviews)) #Creating Corpus of the reviews
negreviews <- VCorpus(VectorSource(negreviews))

clean_corpus <- function(corpus, badwords){  #Function to clean the corpus 
  corpus <- tm_map(corpus, stripWhitespace) #Getting rid of unwanted white space
  corpus <- tm_map(corpus, removePunctuation) #Getting rid of punctuation
  corpus <- tm_map(corpus, content_transformer(tolower)) #making all characters lowercase
  corpus <- tm_map(corpus, removeWords, c("game", "review", "like", "can", "theres", 
                                          "just", "even", "games", "really", "1010", "dont",
                                          "will", "ive", "one", "along", "doesnt", "well", 
                                          "much", "many", "also", "lets", "isnt", "good", "bad", 
                                          "not", "get", "great", "play", "playing", "played", "early",
                                          "access", "b", "u", "®")) #List of unwanted words
  corpus <- tm_map(corpus, removeWords, stopwords("en")) #Removing stopwords
  corpus <- tm_map(corpus, removeNumbers)
  corpus <- tm_map(corpus, removeWords, gamename)
  
  #Removing words that are misspelled
  #Lists larger than 2000 can break the removewords function
  #To fix the problem we do the removal in chunks of 2000 if the list is larger than 2000
  
  if (length(badwords) > 2000){
    s <- 1
    e <- 2000
    
    while (anyNA(badwords[s:e] == FALSE)){ 
      
      corpus <- tm_map(corpus, removeWords, tolower(badwords[s:e]))
      
      if (anyNA(badwords[s:e]) == TRUE) break #If the function is calling values in the list that don't exist, break the loop
      
      s <- s+2000 #Moving to next set of words
      e <- e+2000
      
    }
    
  }
  
  else{
    
    corpus <- tm_map(corpus, removeWords, tolower(badwords)) #If there is less than 2000 words remove them normally
    
  }

  return(corpus) #Returning the cleaned Corpus
  
}

posreviews <- clean_corpus(posreviews, badwords) #Applying the corpus function to the review corpuses
negreviews <- clean_corpus(negreviews, badwords)
```
Here the reviews are converted into volatile corpuses so that the tm package can apply functions to them. A function to tidy and remove certain things from the review text is made. In this function first all extra whitespace and punctuation is removed as well as making all of the characters lowercase. Next a list of words that added no meaning to the result are removed as well as stop words, number characters, and the words in the name of the application or game. Next the list of misspelled words is used to remove those words from the reviews. Because of the way the regular expression for the removewords function is made it can't handle all of words at once so the are removed in groups of 2000 if the list is larger than 2000. The function after it is done returns the tidy corpus.

###Creating the Visualization

```{r, include = TRUE}
find_freq_words <- function(corpus){ #Funcion to find the frequency of each word
  
  BigramTokenizer <-
    function(x)
      unlist(lapply(ngrams(words(x), 2), paste, collapse = " "), use.names = FALSE)
  
  dtm <- TermDocumentMatrix(corpus, control = list(tokenize = BigramTokenizer))
  
  m <- as.matrix(dtm)
  v <- sort(rowSums(m),decreasing=TRUE)
  d <- data.frame(word = names(v),freq=v) #Creating a data frame of the terms and their frequencies
  return(d) #Returning data frame
  
}

posreviewfreq <- find_freq_words(posreviews) #Applying frequency function 
negreviewfreq <- find_freq_words(negreviews)

```
Another function is made to find the most frequent bigrams in each set of reviews. First the list of reviews is tokenized to find the bigrams then a term document matrix is made to find the frequency of each bigram in each set of reviews. The function returns a data frame that can be used for the visualization.

```{r, include = TRUE}
wc_min_freq <- 5 #Variable to adjust the minimum frequency required to be displayed on the word cloud
wc_max_words <- 30 #Variable to adjust the maxium amount of words to be displayed on the word cloud
wc_shape <- "pentagon"

create_wordcloud <- function(wordfreq, wc_min_freq, wc_max_words, wccolor)
{ #Function to create word cloud
  
  set.seed(78)
  ggwordcloud(wordfreq$word, wordfreq$freq, min.freq = wc_min_freq, 
              max.words = wc_max_words, shape = wc_shape, color = wccolor)
  
}

poswordcloud <- create_wordcloud(posreviewfreq, wc_min_freq, wc_max_words, "darkgreen") #Creating word cloud
#poswordcloud #Displaying word cloud

negwordcloud <- create_wordcloud(negreviewfreq, wc_min_freq, wc_max_words, "darkred")
#negwordcloud
```

For the visualization first the minimum word frequency is set. This is to prevent phrases with no meaning from showing up in the word cloud. In this example a phrase would have to have more than 5 occurrences to display in the word cloud. Next the maximum amount of phrases is set. This is to prevent a cluttered word cloud and in this example no more than 30 phrases will apear. The word cloud will display the most common phrases in various sizes, the size of the phrase is based on its frequency. 

###Positive Reviews
```{r echo=FALSE}
poswordcloud
head(posreviewfreq)
```
###Negative Reviews
```{r echo=FALSE}
negwordcloud
head(negreviewfreq)
```
Here are the results for each word cloud and their table of phrase frequencies. When looking at the negative reviews one can notice a lot of talk about personal information and the company Red Shell. The context behind this is that one of the developers for Civilization VI was apart of data controversy where the personal information of users was being shared with thrid party companies. This is only an example of a single game but based on what is shown here, if there is enough data it is possible to get meaningful information from these word clouds.